---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# {rgeedim}

<!-- badges: start -->
[![R-CMD-check](https://github.com/brownag/rgeedim/workflows/R-CMD-check/badge.svg)](https://github.com/brownag/rgeedim/actions)
[![HTML Docs](https://camo.githubusercontent.com/f7ba98e46ecd14313e0e8a05bec3f92ca125b8f36302a5b1679d4a949bccbe31/68747470733a2f2f696d672e736869656c64732e696f2f62616467652f646f63732d48544d4c2d696e666f726d6174696f6e616c)](https://humus.rocks/rgeedim/)
<!-- [![codecov](https://codecov.io/gh/brownag/rgeedim/branch/main/graph/badge.svg?token=BYBKW7PKC3)](https://codecov.io/gh/brownag/rgeedim/) -->
<!-- badges: end -->

{rgeedim} supports search and download of Google Earth Engine imagery with Python `geedim`.

## Installation

You can install the development version of {rgeedim} using {remotes}:

``` r
# install.packages("remotes")
remotes::install_github("brownag/rgeedim")
```

### External Dependencies

You will need Python, and the `geedim` module, installed on your computer in order to successfully use {rgeedim}.

You can install `geedim` with `pip`, for example:

```sh
python -m pip install geedim
```

There are numerous dependencies that may need to be satisfied. Perhaps the most demanding are those for `rasterio`/GDAL. 

If you have trouble compiling the packages yourself on Windows, you can take advantage of the unofficial `pip` wheels (binaries) prepared by by [Christoph Gohlke](https://www.cgohlke.com/): <https://www.lfd.uci.edu/~gohlke/pythonlibs/>. _Note the specific version of Python you are installing for._ Download the desired package/version and then call `pip install your-package.whl`.

## Example

This example shows how to extract a Google Earth Engine asset by name for an arbitrary extent. The coordinates of the bounding box are expressed in WGS84 decimal degrees (`"OGC:CRS84"`).

```{r example}
library(rgeedim)
gd_initialize()

r <- gd_bbox(
  xmin = -120.6032,
  xmax = -120.5377,
  ymin = 38.0807,
  ymax = 38.1043
)

# US NED CHILI (Continuous Heat-Insolation Load Index)
# https://developers.google.com/earth-engine/datasets/catalog/CSP_ERGo_1_0_US_CHILI
x <- gd_image_from_id('CSP/ERGo/1_0/US/CHILI')


res <- gd_download(
  x,
  filename = 'image.tif',
  region = r,
  scale = 10, # scale=10: request ~10m resolution (result in native CRS)
  overwrite = TRUE,
  silent = FALSE
)
```

```{r inspect}
library(terra)

f <- rast(res)
plot(f[[1]])

f
```

```{r include=FALSE}
unlink('image.tif')
```
