---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  eval = !inherits(rgeedim::gd_version(), 'try-error'),
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# {rgeedim}

<!-- badges: start -->
[![R-CMD-check](https://github.com/brownag/rgeedim/workflows/R-CMD-check/badge.svg)](https://github.com/brownag/rgeedim/actions)
[![HTML Docs](https://camo.githubusercontent.com/f7ba98e46ecd14313e0e8a05bec3f92ca125b8f36302a5b1679d4a949bccbe31/68747470733a2f2f696d672e736869656c64732e696f2f62616467652f646f63732d48544d4c2d696e666f726d6174696f6e616c)](https://humus.rocks/rgeedim/)
[![codecov](https://codecov.io/gh/brownag/rgeedim/branch/main/graph/badge.svg?token=BYBKW7PKC3)](https://app.codecov.io/gh/brownag/rgeedim)
[![CRAN status](https://www.r-pkg.org/badges/version-last-release/rgeedim)](https://CRAN.R-project.org/package=rgeedim)
<!-- badges: end -->

{rgeedim} supports search and download of Google Earth Engine imagery with Python module [`geedim`](https://github.com/dugalh/geedim)  by Dugal Harris. This package provides wrapper functions that make it more convenient to use `geedim` from R. 

The [rgeedim manual](https://humus.rocks/rgeedim/) describes the R API. See the [geedim manual]( https://geedim.readthedocs.io/) for more information on Python API and command line interface.

Using `geedim`, images larger than the [`Image.getDownloadURL()` size limit](https://developers.google.com/earth-engine/apidocs/ee-image-getdownloadurl) are split and downloaded as separate tiles, then re-assembled into a single GeoTIFF.

## Installation

{rgeedim} is available on CRAN, and can be installed as follows:

``` r
install.packages("rgeedim")
```

You can install the development version of {rgeedim} using {remotes}:

``` r
if (!require("remotes")) install.packages("remotes")
remotes::install_github("brownag/rgeedim")
```

## Dependencies

You will need Python 3 with the `geedim` module installed to use {rgeedim}.

#### Using `gd_install()`

`gd_install()` provides a simple wrapper around `reticulate::py_install()` that allows you to install dependencies using `pip`, virtual, or Conda environments.

```r
# install with pip (with reticulate)
gd_install()

# install with pip (system() call)
gd_install(system = TRUE)

# use virtual environment with default name "r-reticulate"
gd_install(method = "virtualenv")

# use "conda" environment named "foo"
gd_install(method = "conda", envname = "foo")
```

#### Using `pip`

You can install `geedim` with `pip`, for example:

``` sh
python -m pip install geedim
```

This shell command assumes you have a Python 3 executable named (or aliased) as `python` on your PATH, or the command is being called from within an active Python virtual environment.

#### Using Miniconda

If you do not have a Python environment set up, an option that {reticulate} provides is `reticulate::install_miniconda()`. Once you install Miniconda, you can install packages into a conda environment such as `"r-reticulate"` (default). Customize the `envname` argument to create or add to a specific environment.

``` r
reticulate::install_miniconda()
reticulate::py_install("geedim")
```

### Troubleshooting 

If using Python within RStudio for the first time, you may need to set your default interpreter in _Tools_ >> _Global Options..._ >> _Python_.

If you have trouble compiling dependency packages on Windows, you can take advantage of the unofficial `pip` wheels (binaries) prepared by by [Christoph Gohlke](https://www.cgohlke.com/): <https://www.lfd.uci.edu/~gohlke/pythonlibs/>. _Note the name of the package and the specific version of Python you are installing for._ Download the desired package/version and then call `pip install your-package.whl`.

## How {rgeedim} Works

This example shows how to extract a Google Earth Engine asset by name for an arbitrary extent. The coordinates of the bounding box are expressed in WGS84 decimal degrees (`"OGC:CRS84"`).

```{r example}
library(rgeedim)
```

If this is your first time using any Google Earth Engine tools, authenticate with `gd_authenticate()`. You can pass arguments to use several different authorization methods. Perhaps the easiest to use is `auth_mode="notebook"` in that does not rely on an existing `GOOGLE_APPLICATION_CREDENTIALS` file nor an installation of the `gcloud` CLI tools. However, the other options are better for non-interactive use.

``` r
gd_authenticate(auth_mode = "notebook")
```

In each R session you will need to initialize the Earth Engine library

```{r}
gd_initialize()
```

`gd_bbox()` is a simple function for specifying extents to {rgeedim} functions like `gd_download()`:

``` {r}
r <- gd_bbox(
  xmin = -120.6032,
  xmax = -120.5377,
  ymin = 38.0807,
  ymax = 38.1043
)
```

We will download the US NED CHILI (Continuous Heat-Insolation Load Index) <https://developers.google.com/earth-engine/datasets/catalog/CSP_ERGo_1_0_US_CHILI>. We specify an equal-area coordinate reference system (NAD83 Albers), bilinear resampling, and a resolution of 10 meters.

```{r}
res <- 'CSP/ERGo/1_0/US/CHILI' |>
          gd_image_from_id() |>
          gd_download(
            filename = 'image.tif',
            region = r,
            crs = "EPSG:5070",
            resampling = "bilinear",
            scale = 10, # scale=10: request ~10m resolution
            overwrite = TRUE,
            silent = FALSE
          )
```

We can inspect our results with {terra}. The resulting GeoTIFF has two layers,  `"constant"` and  `"FILL_MASK"`. The former contains the data, and the latter contains a mask reflecting data availability (value = 1 where data are available).

```{r inspect}
library(terra)

f <- rast(res)
f

plot(f[[1]])
```

## Example: Hillshade from DEM

This example demonstrates the download of a section of the USGS NED seamless 10m grid. This DEM is processed with {terra} to calculate some terrain derivatives (slope, aspect) and a hillshade.

```{r}
library(rgeedim)
library(terra)

gd_initialize()

b <- gd_bbox(
  xmin = -120.296,
  xmax = -120.227,
  ymin = 37.9824,
  ymax = 38.0071
)

## hillshade example
# download 10m NED DEM in AEA
x <- "USGS/NED" |>
  gd_image_from_id() |>
  gd_download(
    region = b,
    scale = 10,
    crs = "EPSG:5070",
    resampling = "bilinear",
    filename = "image.tif",
    overwrite = TRUE,
    silent = FALSE
  )
dem <- rast(x)$elevation

# calculate slope, aspect, and hillshade with terra
slp <- terrain(dem, "slope", unit = "radians")
asp <- terrain(dem, "aspect", unit = "radians")
hsd <- shade(slp, asp)

# compare elevation v.s. hillshade
plot(c(dem, hillshade = hsd))
```

## Example: LiDAR Slope Map

This example demonstrates how to access the 1m LiDAR data from USGS. A key step in this process is the use of `gd_composite()` to resample the component images prior to download.

```{r}
library(rgeedim)
library(terra)

# search and download from USGS 1m lidar data collection
gd_initialize()

# wkt->SpatVector->GeoJSON
b <- 'POLYGON((-121.355 37.56,-121.355 37.555,
          -121.35 37.555,-121.35 37.56,
          -121.355 37.56))' |>
  vect(crs = "OGC:CRS84")
r <- gd_region(b)

# note that resampling is done on the images as part of compositing/before download
x <- "USGS/3DEP/1m" |>
  gd_collection_from_name() |>
  gd_search(region = r) |>
  gd_composite(resampling = "bilinear") |>
  gd_download(region = r,
              crs = "EPSG:5070",
              scale = 1,
              filename = "image.tif",
              overwrite = TRUE,
              silent = FALSE) |>
  rast()

# inspect
plot(terra::terrain(x$elevation))
plot(project(b, x), add = TRUE)
```

## Example: Landsat-7 cloud/shadow-free composite

This example demonstrates download of a Landsat-7 cloud/shadow-free composite image. A collection is created from the USGS Landsat 7 Level 2, Collection 2, Tier 1. This example is based on a [tutorial](https://geedim.readthedocs.io/en/latest/examples/l7_composite.html) in the `geedim` manual.

```{r}
library(rgeedim)
library(terra)

gd_initialize()

b <- gd_bbox(
  xmin = -120.296,
  xmax = -120.227,
  ymin = 37.9824,
  ymax = 38.0071
)

## landsat example
# search collection for date range and minimum data fill (85%)
x <- 'LANDSAT/LE07/C02/T1_L2' |>
  gd_collection_from_name() |>
  gd_search(
    start_date = '2020-11-01',
    end_date = '2021-02-28',
    region = b,
    cloudless_portion = 85
  )

# inspect individual image metadata in the collection
gd_properties(x)

# download a single image
y <- gd_properties(x)$id[1] |> 
  gd_image_from_id() |>
  gd_download(
    filename = "image.tif",
    region = b,
    scale = 30,
    crs = 'EPSG:5070',
    dtype = 'uint16',
    overwrite = TRUE,
    silent = FALSE
  )
plot(rast(y)[[1:4]])

# create composite landsat image near December 1st, 2020 and download  
# using q-mosaic method. 
z <- x |> 
  gd_composite(
    method = "q-mosaic",
    date = '2020-12-01'
  ) |>
  gd_download(
    filename = "image.tif",
    region = b,
    scale = 30,
    crs = 'EPSG:5070',
    dtype = 'uint16',
    overwrite = TRUE,
    silent = FALSE
  )
plot(rast(z)[[1:4]])
```

The `"q-mosaic"` method produces a composite largely free of artifacts; this is because it prioritizes pixels with higher distance from clouds.

```{r include=FALSE}
unlink('image.tif')
```
