---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# {rgeedim}

<!-- badges: start -->
[![R-CMD-check](https://github.com/brownag/rgeedim/workflows/R-CMD-check/badge.svg)](https://github.com/brownag/rgeedim/actions)
[![HTML Docs](https://camo.githubusercontent.com/f7ba98e46ecd14313e0e8a05bec3f92ca125b8f36302a5b1679d4a949bccbe31/68747470733a2f2f696d672e736869656c64732e696f2f62616467652f646f63732d48544d4c2d696e666f726d6174696f6e616c)](https://humus.rocks/rgeedim/)
[![codecov](https://codecov.io/gh/brownag/rgeedim/branch/main/graph/badge.svg?token=BYBKW7PKC3)](https://codecov.io/gh/brownag/rgeedim/)
<!-- badges: end -->

{rgeedim} supports search and download of Google Earth Engine imagery with Python module [`geedim`](https://github.com/dugalh/geedim). This package provides wrapper functions that make it more convenient to use `geedim` from R. The [rgeedim manual](https://humus.rocks/rgeedim/) describes the R API. See the [geedim manual]( https://geedim.readthedocs.io/) for more information on Python API and command line interface.

By using `geedim` images larger than the [`Image.getDownloadURL()` size limit](https://developers.google.com/earth-engine/apidocs/ee-image-getdownloadurl) are split and downloaded as separate tiles, then re-assembled into a single GeoTIFF.

## Installation

You can install the development version of {rgeedim} using {remotes}:

``` r
# install.packages("remotes")
remotes::install_github("brownag/rgeedim")
```

## Dependencies

You will need Python 3 with the `geedim` module installed to use {rgeedim}.

#### Using `pip`

You can install `geedim` with `pip`, for example:

``` sh
python -m pip install geedim
```

This shell command assumes you have a Python 3 executable named (or aliased) as `python` on your PATH, or the command is being called from within an active Python virtual environment.

#### Using Miniconda

If you do not have a Python environment set up, an option that {reticulate} provides is `reticulate::install_miniconda()`. Once you install Miniconda, you can install packages into a conda environment such as `"r-reticulate"` (default). Customize the `envname` argument to create or add to a specific environment.

``` r
reticulate::install_miniconda()
reticulate::py_install("geedim")
```

### Troubleshooting 

If using Python within RStudio for the first time, you may need to set your default interpreter in _Tools_ >> _Global Options..._ >> _Python_.

If you have trouble compiling dependency packages on Windows, you can take advantage of the unofficial `pip` wheels (binaries) prepared by by [Christoph Gohlke](https://www.cgohlke.com/): <https://www.lfd.uci.edu/~gohlke/pythonlibs/>. _Note the name of the package and the specific version of Python you are installing for._ Download the desired package/version and then call `pip install your-package.whl`.

## Example

This example shows how to extract a Google Earth Engine asset by name for an arbitrary extent. The coordinates of the bounding box are expressed in WGS84 decimal degrees (`"OGC:CRS84"`).

```{r example}
library(rgeedim)
```

If this is your first time using any Google Earth Engine tools, authenticate with `gd_authenticate()`. You can pass arguments to use several different authorization methods. Perhaps the easiest to use is `auth_mode="notebook"` in that does not rely on an existing `GOOGLE_APPLICATION_CREDENTIALS` file nor an installation of the `gcloud` CLI tools. However, the other options are better for non-interactive use.

``` r
gd_authenticate(auth_mode = "notebook")
```

In each R session you will need to initialize the Earth Engine library

```{r}
gd_initialize()
```

`gd_bbox()` is a simple function for specifying extents to {rgeedim} functions like `gd_download()`:

``` {r}
r <- gd_bbox(
  xmin = -120.6032,
  xmax = -120.5377,
  ymin = 38.0807,
  ymax = 38.1043
)

# US NED CHILI (Continuous Heat-Insolation Load Index)
# https://developers.google.com/earth-engine/datasets/catalog/CSP_ERGo_1_0_US_CHILI
res <- 'CSP/ERGo/1_0/US/CHILI' |>
  gd_image_from_id() |>
  gd_download(
    filename = 'image.tif',
    region = r,
    crs = "EPSG:5070",
    resampling = "bilinear",
    scale = 10, # scale=10: request ~10m resolution (result in native CRS)
    overwrite = TRUE,
    silent = FALSE
  )
```

```{r inspect}
library(terra)

f <- rast(res)
plot(f[[1]])

f
```

```{r include=FALSE}
unlink('image.tif')
```
