---
title: "`rgeedim` Demo"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{`rgeedim` Demo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  # only run vignette if modules available
  collapse = TRUE,
  comment = "#>"
)
```

This vignette shows how to extract a Google Earth Engine asset by name for an arbitrary extent and visualize it in R.

First, we load {rgeedim} and call `gd_initialize()`. This is a wrapper function around `geedim.Initialize()` that must be run before using the Python Google Earth Engine API.

```{r setup}
library(rgeedim)
gd_initialize()
```

Perhaps the _simplest_ way to specify the target extent is using the xmin/xmax/ymin/ymax arguments to `gd_bbox()`. This function returns a Python object equivalent to _GeoJSON_, which is interchangeably represented as a simple `list` object in **R** using {reticulate}. 

## Determine Target Region

```{r bbox}
r <- gd_bbox(
  xmin = -121,
  xmax = -120.5,
  ymin = 38.5,
  ymax = 39
)
```

As is standard for _GeoJSON_, coordinates of the bounding box are expressed in WGS84 decimal degrees (`"OGC:CRS84"`). Note that longitude, latitude (X, Y) coordinate pair order is implied.

## Access by ID

We can find IDs of assets of interest using the Google Earth Engine data catalog: <https://developers.google.com/earth-engine/datasets/catalog>

To obtain an R object reference to the asset we pass the `"id"` to `gd_image_from_id()`. For example here we use [Global SRTM Topographic Diversity](https://developers.google.com/earth-engine/datasets/catalog/CSP_ERGo_1_0_Global_SRTM_topoDiversity):

```{r image-from-id}
x <- gd_image_from_id('CSP/ERGo/1_0/Global/SRTM_topoDiversity')
```

Depending on the resource `gd_image_from_id()` will either return a `geedim.mask.MaskedImage` or `geedim.mask.MaskedCollection` object

Now, we pass the image result to `gd_download()`. We can specify output `filename` and target area as `region` arguments. See `gd_bbox()` for examples of making a region argument from bounding coordinates or a {terra} `SpatExtent` object.

Other options that can be passed to the `BaseImage.download()` method include `scale` which allows warping of the result to a target resolution. Try modifying this example to use `scale=90` (~native SRTM resolution):

```{r download}
img <- gd_download(x, filename = 'image.tif',
                   region = r, scale = 900,
                   overwrite = TRUE, silent = FALSE
                  )
```

`gd_download()` (invisibly) returns the `filename` on successful download, which helps to "pipe" into functions that might read the result.

So we can use the {terra} ` rast()` function to read the GeoTIFF `gd_download()` result. 

```{r terrarast, fig.align='center', fig.width=4}
library(terra)
f <- rast(img)
```

```{r inspect}
par(mar=c(1,1,1,1))
plot(f[[1]])

# inspect object
f
```

```{r include=FALSE}
unlink("image.tif")
```
